---
title: "A Gentle Guide to Tidy Statistics in R"
date: 03-16-2018
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


### Load packages
Using the `library` function load your packages eg `library(package_name)`

```{r, message=F, warning=FALSE, echo = T}
# Load libraries
library(tidyverse) # plotting, cleaning, etc
library(broom) # nice statistical outputs
library(knitr) # some tables and R Markdown
library(readxl) # read in Excel files
library(here) # nice file paths
```

### Step 1 Load the data!

We have to tell R what format of data we are trying to read in (.csv, .xls, .tsv, etc)

```{r, eval = TRUE}
#### Read excel file
raw_df <- read_xlsx("ad_treatment.xlsx")
glimpse(raw_df)
```

***
  

```{r}
ggplot(data = raw_df, aes(x = mmse)) + 
 geom_density(fill = "blue")
```
  
***

```{r}
raw_df %>% 
  summarize(
    min = min(mmse), 
    max = max(mmse)
    )
```

```{r}
raw_df %>% 
  group_by(health_status) %>% 
  summarize(
    min = min(mmse), 
    median = median(mmse),
    max = max(mmse)
    )
```

***
  

```{r, echo = T}
raw_df %>% 
  group_by(drug_treatment, health_status) %>% 
  count()
```

### Data exploration of dependent variable

Before running our summary statistics we can actually visualize the range, central tendency and quartiles via a geom_boxplot call.

```{r, echo = T, fig.align='center'}
ggplot(data = raw_df, # add the data
       aes(x = drug_treatment, y = mmse, # set x, y coordinates
           color = drug_treatment)) +    # color by treatment
  geom_boxplot() +
  facet_grid(~health_status) # create panes base on health status
```

***

### Summary Statistics


```{r, echo = F}
dplyr::glimpse(raw_df)
```

***

```{r, echo = F}
sum_df <- raw_df %>% 
            mutate(
              sex = factor(sex,
                              levels = c("0", "1"),
                  labels = c("Male", "Female")),
              drug_treatment =  factor(drug_treatment, 
                  levels = c("Placebo", "Low dose", "High Dose")),
              health_status = factor(health_status, 
                  levels = c("Healthy", "Alzheimer's"))
              )

glimpse(sum_df)
```



```{r, echo = T}
sum_df <- sum_df %>%   
  group_by(sex, health_status, drug_treatment) %>%  
  summarize(mmse_mean = mean(mmse),   
            mmse_se = sd(mmse)/sqrt(n()),
            n_samples = n()) %>%
  ungroup() # ungrouping variable is a good habit to prevent errors
```
  
***  
  
Now we have a nicely formatted dataframe that can be saved to Excel, or used in graphing. We need to indicate what data we are writing (sum_df) and what we want the resulting file to be named (“adx37_sum_stats.csv”).

```{r, eval = F, echo = T}
# code to save the table into a .csv Excel file
write_csv(sum_df, "adx37_sum_stats.csv")
```
  
***  
  
### Summary graph

By calling a ggplot function we can generate a preliminary summary graph.

```{r, echo = T}
ggplot(data = sum_df, # add the data
       aes(x = drug_treatment,  #set x, y coordinates
           y = mmse_mean,
           group = drug_treatment,  # group by treatment
           color = drug_treatment)) +    # color by treatment
  geom_point(size = 3) + 
  facet_grid(sex~health_status) # create facets by sex and status
```

### The ANOVA finally!


```{r}
stats_df <- raw_df %>% # start with data
   mutate(drug_treatment = factor(drug_treatment, 
                                  levels = c("Placebo", "Low dose", "High Dose")),
         sex = factor(sex, 
                      levels = c("0", "1"),
                      labels = c("Male", "Female")),
         health_status = factor(health_status, 
                                levels = c("Healthy", "Alzheimer's")))
glimpse(stats_df)
```
  
***  
  

```{r, eval = F}
aov(dependent_variable ~ independent variable, data = data_df)
```

```{r, eval = F}
# this would give ONLY main effects
ad_aov <- aov(mmse ~ sex + drug_treatment + health_status, data = stats_df)
```

```{r}
# this gives main effects AND interactions
ad_aov <- aov(mmse ~ sex * drug_treatment * health_status, 
        data = stats_df)
```


```{r, eval = F}
# this throws an error because we shouldn't use commas in between!
ad_aov <- aov(mmse ~ sex, drug_treatment, health_status, data = stats_df)
```
  
***

  
```{r}
# look at effects and interactions
summary(ad_aov)
```


```{r, eval = F}
# this extracts ANOVA output into a nice tidy dataframe
tidy_ad_aov <- tidy(ad_aov)
# which we can save to Excel
write_csv(tidy_ad_aov, "ad_aov.csv")
```

### Post-hocs

```{r}
# call and save the pair.t.test
ad_pairwise <- pairwise.t.test(stats_df$mmse,    
                               stats_df$sex:stats_df$drug_treatment:stats_df$health_status, 
                               p.adj = "bonferroni")
```


```{r}
# tidy the post hoc
tidy_ad_pairwise <- broom::tidy(ad_pairwise) %>% 
  mutate(p.value = round(p.value, 5))
# look at the comparisons and p-values
head(tidy_ad_pairwise)
```

  
***  


```{r, echo = F}
TukeyHSD(ad_aov, which = 'sex:drug_treatment:health_status') %>% 
  tidy() %>% 
  head() %>% 
  kable()
```

### Publication Graph

```{r}
tribble(
  ~colA, ~colB,
  "a",   1,
  "b",   2,
  "c",   3
)
```

And here is our actual code for making the custom dataframe.

```{r}
# make the dataframe with specific points of interest to add *
sig_df <- tribble(
  ~drug_treatment, ~ health_status, ~sex, ~mmse_mean,
  "Low dose", "Alzheimer's", "Male", 17,
  "High Dose", "Alzheimer's", "Male", 25,
  "Low dose", "Alzheimer's", "Female", 18, 
  "High Dose", "Alzheimer's", "Female", 24
  )
# convert the variables to factors again :)
sig_df <- sig_df %>% 
  mutate(drug_treatment = factor(drug_treatment, 
               levels = c("Placebo", "Low dose", "High Dose")),
         sex = factor(sex, 
               levels = c("Male", "Female")),
         health_status = factor(health_status, 
               levels = c("Healthy", "Alzheimer's")))
```

And the final graph!

```{r, echo = F, fig.height=6, fig.width=7}
g1 <- ggplot(data = sum_df, 
       aes(x = drug_treatment, y = mmse_mean, fill = drug_treatment,  
           group = drug_treatment)) +
  geom_errorbar(aes(ymin = mmse_mean - mmse_se, 
                    ymax = mmse_mean + mmse_se), width = 0.5) +
  geom_bar(color = "black", stat = "identity", width = 0.7) +
  
  facet_grid(sex~health_status) +
  theme_bw() +
  scale_fill_manual(values = c("white", "grey", "black")) +
  theme(legend.position = "NULL",
        legend.title = element_blank(),
        axis.title = element_text(size = 20),
        legend.background = element_blank(),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 12)) +
  geom_text(data = sig_df, label = "*", size = 8) +
  labs(x = "\nDrug Treatment", 
       y = "Cognitive Function (MMSE)\n",
       caption = "\nFigure 1. Effect of novel drug treatment AD-x37 on cognitive function in healthy and demented elderly adults. \nn = 100/treatment group (total n = 600), * indicates significance at p < 0.001")
g1
```

```{r, eval = F}
# save the graph!
ggsave("ad_publication_graph.png", g1, height = 7, width = 8, units = "in", dpi = 500)
```


```{r}

library(ggridges)
stats_df %>% 
  group_by(sex, drug_treatment, health_status) %>% 
  mutate(median_line = median(mmse)) %>% 
  ggplot(aes(x = mmse, fill = drug_treatment)) +
  geom_density() +
    geom_vline(aes(xintercept=median_line, color = drug_treatment)) +
  facet_grid(sex~health_status)
```

```{r}
stats_df %>% 
  ggplot(aes(y = drug_treatment, x = mmse, fill = drug_treatment)) +
  geom_density_ridges(quantile_lines = TRUE, quantiles = 2) +
  facet_grid(sex~health_status)
```

