---
title: "class_04"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(socviz)

#load friends data
full_data <- read_csv("data/full_data.csv")


#create list for filter
target <- c ("Rachel Green", "Ross Geller", "Joey Tribbiani", "Chandler Bing", "Monica Geller", "Phoebe Buffay")

#filter rows for only the main characters and drop rows with NAs
fulldata_main <- full_data%>%
  filter(speaker %in% target)%>%
  drop_na()

#note the new object that contains less rows
```

## Class Assessment (Where are you in your R journey?)

These are meant to help you assess where you are in your R journey. If you can answer both of these questions without help, then you are a in a great place. If you struggle with these, practice some more. I recommend the primers you worked on last week or the exercises at the end of the chapters in R4.DS. 

```{r Question 1}

#look at the basics of the midwest dataset
summary(midwest)

#building a histogram based on the total population by county
ggplot(data = midwest, mapping = aes(x = poptotal))+
  geom_histogram(bins = 100)+
  annotate("text", x = 4500000, y = 25, label = "Which County Is This?")+
  labs(title = "Histogram of Midwestern Counties By Population", 
       x = "Population", y = "Number of Counties")

```


```{r Quesetion 2}
#build a basic dataset
ages <- tribble(
  ~speaker,      ~age,
  "Rachel",         24,
  "Phoebe",         27,
  "Ross",           26,
  "Monica",         24,
  "Chandler",       26,
  "Joey",           26
)

#create a new object that holds the mean of all the character's ages
c_avg <- mean(ages$age)

#create a bar chart that visualizes the character's age and mean
ggplot(data = ages) +
  geom_bar(mapping = aes(x = speaker, 
                         y = age, 
                         fill = speaker), 
           stat = "identity")+
  geom_hline(aes(yintercept = c_avg))+
  labs(title = "How old were the Friends characters?", 
       x = "Character", y = "Age (in years)")+
  guides(fill=FALSE)
```


```{r create bar graph}
#default stat transformation fro geom_bar() is stat_count()
ggplot(data = fulldata_main)+
  geom_bar(mapping = aes(x = speaker))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# no stat transformation        
ages <- tribble(
  ~speaker,      ~age,
  "Rachel",         23,
  "Phoebe",         28,
  "Ross",           31,
  "Monica",         24,
  "Chandler",       24,
  "Joey",           26
)

c_avg <- mean(ages$age)

ggplot(data = ages) +
  geom_bar(mapping = aes(x = speaker, 
                         y = age, 
                         fill = speaker), 
           stat = "identity")+
  geom_hline(aes(yintercept = c_avg))+
  labs(title = "How old were the Friends characters?", 
       x = "Character", y = "Age (in years)")+
  guides(fill=FALSE)


```

## EDA

```{r}
#stacked bar graph
ggplot(data = fulldata_main, mapping = aes(x = emotion, fill = speaker)) +
  geom_bar()

#emotion vs speaker faceted
ggplot(data = fulldata_main, 
       mapping = aes(x = emotion, y = stat(prop), 
                     group = 1, fill = speaker))+
  geom_bar()+
  facet_wrap(~speaker)+
  coord_flip()

#speaker vs emotion faceted
ggplot(data = fulldata_main, 
       mapping = aes(x = speaker, y = stat(prop), 
                     group = 1, fill = emotion))+
  geom_bar()+
  facet_wrap(~emotion)+
  coord_flip()

#create list of emotions that we want to highlight
select_emotions <- c("Sad","Mad","Scared")

#visualize selected emotions and clean up graph
fulldata_main%>%
  filter(emotion%in%select_emotions)%>%
  ggplot(mapping = aes(x = speaker, y = stat(prop), 
                       group = 1, fill = emotion))+
  geom_bar()+
  facet_wrap(~emotion)+
  guides(fill = FALSE)+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
## Data Transformations with dplyr
```{r}
#simply but verbose way of filtering data and dropping NAs
full_data_new <- filter(full_data, speaker %in% target)
full_data_new_na <- drop_na(full_data_new)

#better way of filtering with pipes
full_data_new_better <- full_data%>%
  filter(speaker %in% target)%>%
  drop_na()

#creating new aggregated (n) variables 
emotion_by_speaker <- full_data_new_na %>%
  group_by(speaker,emotion)%>%
  summarize(N=n())%>%
  mutate(freq = N / sum(N), pct = round((freq*100),1))

speaker_by_emotion <- full_data_new_na %>%
  group_by(emotion,speaker)%>%
  summarize(N=n())%>%
  mutate(freq = N / sum(N), pct = round((freq*100),1))
```

##Organ Donor EDA

How does organ donation differ across countries?
Does the type of government impact organ donation trends?

```{r plot continuous variables versus categorical}
str(organdata)

ggplot(data = organdata, mapping = aes(x = country, y = donors))+
  geom_boxplot()

ggplot(data = organdata, mapping = aes(x = country, y = donors))+
  geom_boxplot() +
  coord_flip()

ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm=TRUE), y = donors))+
  geom_boxplot()+
  labs(x = NULL)+
  coord_flip()

ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm=TRUE), 
                                       y = donors,
                                       fill = world))+
  geom_boxplot()+
  labs(x = NULL)+
  coord_flip()+
  theme(legend.position = "top")

ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm=TRUE), 
                                       y = donors,
                                       color = world))+
  geom_point()+
  labs(x = NULL)+
  coord_flip()+
  theme(legend.position = "top")

ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm=TRUE), 
                                       y = donors,
                                       color = world))+
  geom_jitter()+
  labs(x = NULL)+
  coord_flip()+
  theme(legend.position = "top")

ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm=TRUE), 
                                       y = donors,
                                       color = world))+
  geom_jitter(position = position_jitter(width = 0.15))+
  labs(x = NULL)+
  coord_flip()+
  theme(legend.position = "top")
```

```{r plotting means and SD}
#verbose way (not comprehensive - just example of first 3)
by_country_verbose <-  organdata %>% group_by(consent_law, country)%>%
  summarise(donors_mean = mean(donors, na.rm = TRUE),
            donors_sd = sd(donors, na.rm = TRUE),
            gdp_mean = mean(gdp, na.rm = TRUE))

#better way but requires summarize_if() 
by_country <- organdata %>%
  group_by(consent_law, country) %>%
  summarize_if(is.numeric,
               list(~ mean(., na.rm = TRUE),
                    ~ sd(., na.rm = TRUE))) %>%
  ungroup()
by_country

ggplot(data = by_country,
       mapping = aes(x = donors_mean,
                     y = reorder(country, donors_mean),
                     color = consent_law))+
  geom_point(size = 3) +
  labs(x = "Donor Procurement Rate",
       y = "", 
       color = "Consent Law")+
  theme(legend.position = "top")


ggplot(data = by_country, mapping = aes(x = reorder(country, donors_mean),y = donors_mean))+
         geom_pointrange(mapping = aes(ymin = donors_mean - donors_sd, 
                                       ymax = donors_mean + donors_sd,
                                       color = consent_law))+
         labs(x = "", y="Donor Procurement Rate")+
         coord_flip()+theme(legend.position = "top")
```


```{r ploting text on graphs}
ggplot(data = by_country,
       mapping = aes(x = roads_mean, y = donors_mean))+
  geom_point()

ggplot(data = by_country,
       mapping = aes(x = roads_mean, y = donors_mean))+
  geom_point() +
  geom_text(mapping = aes(label = country))

ggplot(data = by_country,
       mapping = aes(x = roads_mean, y = donors_mean))+
  geom_point() +
  geom_text(mapping = aes(label = country), hjust = 0)

ggplot(data = by_country,
       mapping = aes(x = roads_mean, y = donors_mean))+
  geom_point() +
  geom_text(mapping = aes(x = roads_mean + 1, label = country), hjust = 0)

ggplot(data = by_country,
       mapping = aes(x = roads_mean, y = donors_mean))+
  geom_point() +
  geom_text(mapping = aes(label = country), nudge_x = 5)
```
```{r working with scales}
by_country_world <- organdata %>%
  group_by(world, country, donors) %>%
  summarize_if(is.numeric,
               list(~ mean(., na.rm = TRUE),
                    ~ sd(., na.rm = TRUE))) %>%
  ungroup()
by_country_world

d <- ggplot(data = by_country_world,
            mapping = aes(x = roads_mean,
                          y = donors, color = world))
d + geom_point() +
  scale_x_log10() + scale_y_continuous(breaks = c(5, 15, 25),
                                       labels = c("Five", "Fifteen", "Twenty Five"))

e <- ggplot(data = by_country_world, mapping = aes(x = roads_mean, y = donors, color = world))
e + geom_point() + scale_color_discrete(labels = c("Corporatist", "Liberal","Social Democratic", "Unclassified")) +
  labs(x = "Road Deaths", y = "Donor Procurement",color = "Welfare State")+
  guides(color = FALSE)
```

