---
title: "EDA"
author: "Andrew Bell"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
```

## *Learning Objectives*

-   Become familiar with data transformations
-   Understand how to deal with na's
-   Learn how manipulate labels and values in ggplot 
-   

## Data Transformations

In essence, most everything you do in R involves **objects**. **Functions** take **arguments** (some of which are objects) and produce outputs (which you can assign to a new object).

Data frames are made up of multiple lists. You can create dataframes using the data_frame() function.

```{r}

```

-   

## How to Deal with NA's

We want to be able to make visualizations from data - so we need a function that outputs a plot based on the objects and arguments we give it. Enter the **ggplot()** function. Below is most basic syntax (structure of code re

## Learn how manipulate labels and values in ggplot

I want you to now create your own visualization that explores the how the shows popularity changed over time.
