---
title: "class_04_2021 - learning activity"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data and explore

Explore the opiates dataset (found in socviz). What are the variables? Observations? Start thinking about what you might want to explore. 

Considering sharing one of your groups projects with your group so you can work off a single RMD file. 

```{r Setup }
library(socviz)

str(opiates)

opiates
```

## Exploratory Question

Develop an exploratory question that requires creating new aggregated variable(s) and write it out below. 

Which states are most impacted by the opiate crisis and has the crisis gotten worse over time? 

## Develop code to create your new variables

Write our the code required to create your new aggregated variables
```{r Aggregation_Summarize}
#creating a new variable that gives us death rate per 100k citizen
opiates_withper100k <- opiates %>%
  mutate(death_per100k = deaths/(population/100000))

#creating new summary table with mean death rate per 100k citizen across time
opiates_meanper100k <- opiates_withper100k %>%
  group_by(state,division_name)%>%
  summarise(mean_per100k = mean(death_per100k))

```

##Create visualizations that address your exploratory question
Create as many visualization as you need to properly address your question. 

```{r Visualization 1}
across_mean <- mean(opiates_meanper100k$mean_per100k)

ggplot(opiates_meanper100k, aes(x = mean_per100k, 
                                y = reorder(state,mean_per100k,na.rm = TRUE)))+
  geom_point()+
  geom_vline(aes(xintercept = across_mean))+
  theme_classic()
  
```

```{r Visualization 2}

```

