---
title: ''
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Assignment 2

For this assignment I'd like for you to explore the power of ggplot while also applying your data visualization critique knowledge to your new visualizations

### Step 1 (1pt) 

Load tidyverse package. 
Load gapminder package 
Assign the data in the gapminder package to an object "data"

```{r data}
#your code below


```

### Step 2 (2pts)

Explore the variables and dimensions of your newly created object using the head() and dim() functions

```{r get to know your data}
#your code below


```

### Step 3 (2pts)

Create a data visualization that explores the relationship between year and life expectancy in the gapminder data

```{r data visualization}
#your code below


```

### Step 4 (2pts)
Evaluate your data visualization based on the 5 principles of an effective data visualizations we talked about in class on 9/4. Make sure to use *markdown* effectively. 

---your evaluation here---

### Step 5 (3pts)
Create a new data visualization that adds an additional variable to your visualization (maps a new variable to a new aesthetic). Describe the resulting visualization (3 sentences or less)

```{r data visualization number 2}
#your code below


```